State of the Practice of Lattice Boltzmann Method Software Development

### 2021-06-03 StateOfPractice/Peter-Notes/Report/Report.tex
2nd edits to results section of LBM Report

Sixteen of the software packages have a requirement specification or explicitly reference domain theory. This information is often found in a user manual, on a web page, or is mentioned in publications by the developers. In the latter case the user may need to spend significant time to find this information.

Interviews with developers confirmed that many of the software packages were developed by domain experts with backgrounds in physics, mathematics and mechanical engineering. Many did not have formal software engineering education. Some development teams include computer scientists. Despite a lack of visible domain documentation and a resulting lower surface correctness and verifiability score, many software packages do have a strong domain background.

Document generation tools are explicitly used by 12 software packages. Sphinx is used by 8 of them, and Doxygen is used by 7. Some of the packages use both.

Tutorials are available for all but 5 of the software packages. Generally they are linearly written and easy to follow. However, only 8 tutorials provided an expected output and it is not possible to verify the correctness of the output of the software packages that are missing this key information. In this case the user cannot be sure of correctness and may need to assume it if there are no immediately visible errors.

Unit tests are only explicitly available for one of the software packages. Code modularization of most packages allows for a user to create tests with varying degrees of effort. These tests allow developers and users to verify the correctness of fragments of the source code, and in doing so better assess the correctness of the entire package.

The documentation or presence of continuous integration tools and techniques alludes to a more refined development process where faults are isolated and better recognized. On the surface only 2 of the packages mentioned applying the practice of continuous integration in their development process.

Interviews with the developers of software packages suggest a more frequent use of both unit testing and continuous integration in the development processes than initially surveyed. For example, OpenLB, pyLBM, and TCLB use such methods during development despite it not being explicitly clear from a surface analysis.

@@

Figure \ref{Fig_CorrectnessVerifiability} shows the surface correctness and verifiability ranking of the software packages using AHP. Software packages with a higher score tend to have a visible requirements specification or references to theory documentation. They also explicitly use at least one document generation tool that builds confidence of correctness. The top ranked software packages all include an easy to follow getting started tutorial in their documentation, and most provide expected output for these tutorials. Only the top ranked package, Ludwig, provided unit testing. It and the second ranked package, ESPResSo, explicitly incorporated continuous integration in the development process. During developer interviews it was discovered that some other packages also use continuous integration and unit testing, despite not being explicitly stated in documentation. Such packages, including OpenLB, pyLBM, and TCLB, have higher true correctness and verificablity than is noted from surface analysis.

The inclusion of requirements specification and theory documentation greatly benefits the correctness and verifiability of software packages. The use of document generation tools can help build confidence in correctness. The addition of easy to follow tutorials, along with expected output, further helps users verify the software and have confidence in its correctness. Unit testing documentation and capability, as well as the use of continuous integration tools and techniques such as Bamboo, Jenkins, and Travis CI, help verify correctness. Several interviewed developers alluded to difficulty with testing large numbers of features, and some even manually tested output. The use of well defined unit testing tools could decrease the time spent testing some feature.

@@

Errors occurred when installing 16 of the software packages, and every instance prompted a message. Error messages included unrecognized commands (even when following the installation guide), missing links, missing dependencies, syntax errors in code files, and vague output. Several automatic installation processes could not find and load dependency packages. The installation files pointed to outdated external repositories. Seven of the installations were clearly recovered and one of the installations was assumed to be recovered due to the absence of installation verification. The installation of eight of the software packages could not be recovered. Most of these installations could not reach a dependency repository, had an incompatible error in their core installation files, or prompted vague error messages.

Of the 13 software packages that installed correctly and had tutorials, 4 broke during tutorial testing. All four error resulted in an error message being displayed. One error was due to a missing tutorial dependency, another was due to an invalid command despite following the tutorial, and the final two errors were execution errors. Of the 4 broken tutorial instances, only the one that was missing a dependency was recoverable.

@@

Figure \ref{Fig_Reliability} shows the surface reliability ranking of the software packages using AHP. Software packages with a high score did not break during installation, or the installation was recoverable. All of the top 5 ranked packages had tutorials. One of these packages, pyLBM, broke during tutorial testing, but a descriptive error message helped in recovery.

Overall, lower ranked software packages were lacking clear documentation, testing or tutorial examples, and descriptive error messages, and had broken automatic dependencies. Thus, regarding surface reliability, software packages would benefit from clear up-to-date installation and tutorial documentation, and descriptive error messages.

@@

The software packages were tested to handle unexpected input, including incorrect data types, empty input, and missing files or links. Success predicated on a reasonable response from the system, including appropriate error messages and an absence of unrecoverable system failures.

@@

Figure \ref{Fig_Robustness} shows the surface robustness ranking of the software packages using AHP. Software packages with a high score behaved reasonably in response to unexpected input as described above. All of the software packages that installed correctly passed this test. They output descriptive error messages or did not crash. Software packages with a lower surface robustness score had not installed correctly, so their score here may not be a true reflection of runtime robustness. Similarly, all software packages that installed correctly and require plain text input files handled an unexpected change, a replacement of new lines with carriage returns, to these input files correctly.

@@

Each software package was not quantitatively measured for surface performance.
Although the software packages all apply Lattice Boltzmann methods to solve scientific computing problems, the packages focus on varied domains, with varying parameters, and are technically different from each other. Due to this, a comparison of performance is not appropriate. Instead, in this project we looked through each software package's artifacts for evidence that performance was considered. The artifacts of 17 software packages mentioned the use of some sort of parallelization for execution. This included GPU processing and the CUDA parallel computing platform. GPUs provide superior processing power and speed compared to CPUs, and are often used for scientific computing when a large amount of data is involved. The software package TCLB is implemented in a highly efficient multi-GPU code to achieve performance suitable for model optimization \citep{rutkowski2020open}. In the Ludwig package, a so-called mixed mode approach is used where fine-grained parallelism is implemented on the GPU, and MPI is used for even larger scale parallelism \citep{gray2013ludwig}. While some software packages required CUDA and GPU processing, some had the option of using either the GPU or the CPU. The packages that require GPU and CUDA have better performance at the expense of installability and surface reliability.

>>>>

Sixteen of the software packages include a requirements specification artifact or explicitly reference domain theory. This artifact is often found within a user manual, on a web page, or is mentioned in related publications. In the latter case the user may need to spend significant time to find this information.

Interviews with developers confirmed that many of the software packages are developed by domain experts with backgrounds in physics, mathematics and mechanical engineering. Many of these developers do not have formal software engineering education. Some of the development teams include computer scientists. Despite a lack of visible domain documentation and a resulting lower surface correctness and verifiability score, many software packages do have a strong domain background on account of their development team.

Document generation tools are explicitly used by twelve software packages. Sphinx is used by eight of them, and Doxygen is used by seven. Several of the packages use both.

Tutorials are available for all but five of the software packages. Generally they are linearly written and easy to follow. However, only eight tutorials provide an expected output. It is not possible to verify the correctness of the output of the software packages that are missing this key information. In these cases the user may need to assume correctness if there are no visible errors.

Unit tests are only explicitly available for one of the software packages. Code modularization of most packages allow for users to create tests with varying degrees of effort. These tests allow developers and users to verify the correctness of fragments of the source code, and in doing so better assess the correctness of the entire package.

The documentation or presence of continuous integration tools and techniques alludes to a more refined development process where faults are isolated and better recognized. Only two of the packages mentioned applying the practice of continuous integration in their development process.

Interviews with the developers of software packages suggested a more frequent use of both unit testing and continuous integration in the development processes than initially surveyed. For example, OpenLB, pyLBM, and TCLB use such methods during development despite it not being explicitly clear from a surface analysis.

@@

Figure \ref{Fig_CorrectnessVerifiability} shows the surface correctness and verifiability ranking of the software packages using AHP. Software packages with a higher score tend to have a visible requirements specification or references to theory documentation. They also explicitly use at least one document generation tool that builds confidence of correctness. The top ranked software packages all include an easy to follow getting started tutorial, and most of these include expected output. Only the top ranked package, Ludwig, provided unit testing. It and the second ranked package, ESPResSo, explicitly incorporated continuous integration in the development process. During developer interviews it was discovered that some other packages also use continuous integration and unit testing, despite not being explicitly stated in documentation. The correctness and verificablity of such packages is not measured well using surface analysis.

The inclusion of requirements specification and theory documentation greatly benefits the correctness and verifiability of software packages. The use of document generation tools can help build confidence in correctness. The addition of easy to follow tutorials further helps users verify the software and have confidence in its correctness. Unit testing documentation and capability, as well as the use of continuous integration tools and techniques such as Bamboo, Jenkins, and Travis CI, help verify correctness. Several interviewed developers alluded to difficulty with testing large numbers of features, and some even manually tested program output. The use of well defined unit testing tools could decrease the time spent testing some feature.

@@

Errors occurred when installing sixteen of the software packages. Every instance prompted an error message. These messages indicated unrecognized commands (even when following the installation guide), missing links, missing dependencies, syntax errors in code files. In some instances the error messages were vague. Several automatic installation processes could not find and load dependencies. In these instances the installation tried to access outdated external repositories. Seven of the installations were recovered and verified, and one of the installations was assumed to be recovered due to the absence of any way to verify it. The installation of eight of the software packages could not be recovered. Most of these broken installations could not find external dependencies, encountered system incompatibilities, or displayed vague error messages.

Of the thirteen software packages that installed correctly and also have tutorials, four broke during tutorial testing. All of these instances resulted in an error message being displayed. One error was due to a missing tutorial dependency, another was due to an invalid command despite following the tutorial, and the final two errors were vague execution errors. Of the four broken tutorial instances, only the one that was missing a dependency was recoverable.

@@

Figure \ref{Fig_Reliability} shows the surface reliability ranking of the software packages using AHP. Software packages with a high score either did not break during installation, or the broken installation was recoverable. All of the top five ranked packages have tutorials. One of these packages, pyLBM, broke during tutorial testing, but a descriptive error message helped in recovery.

Overall, lower ranked software packages are lacking clear documentation, testing or tutorial examples, and descriptive error messages, and have broken dependencies. Thus, regarding surface reliability, software packages would benefit from clear up-to-date installation and tutorial documentation, and descriptive error messages.

@@

The software packages were tested for handling unexpected input, including incorrect data types, empty input, and missing files or links. Success predicated on a reasonable response from the system, including appropriate error messages and an absence of unrecoverable system failures.

@@

Figure \ref{Fig_Robustness} shows the surface robustness ranking of the software packages using AHP. Software packages with a high score behaved reasonably in response to unexpected input as described above. All of the software packages that installed correctly passed this test. They output descriptive error messages or did not crash. Software packages with a lower surface robustness score had not installed correctly, so their robustness score may not be a true reflection of runtime robustness. Similarly, all software packages that installed correctly and require plain text input files correctly handled an unexpected change to these input files, including a replacement of new lines with carriage returns.

@@

The software packages were not quantitatively measured for surface performance.
Although the software packages all apply LBM to solve scientific computing problems, the packages focus on varied domains, with varying parameters, and are technically different from each other. Due to this, a comparison of performance is not appropriate. In this project we instead looked through each software package's artifacts for evidence that performance was considered. The artifacts of seventeen software packages mentioned the use of some sort of parallelization tool or technique. This included GPU processing and the CUDA parallel computing platform. GPUs provide superior processing power and speed compared to CPUs, and are often used for scientific computing when a large amount of data is involved. The software package TCLB is implemented in a highly efficient multi-GPU code to achieve performance suitable for model optimization \citep{rutkowski2020open}. In the Ludwig package, a so-called mixed mode approach is used where fine-grained parallelism is implemented on the GPU, and MPI is used for even larger scale parallelism \citep{gray2013ludwig}. While some software packages require CUDA and GPU processing, some have the option of using either the GPU or the CPU. The packages that require GPU and CUDA have better performance at the expense of installability and surface reliability.
